# Chapter 2 - Applied

# 9

## a)

Load in the college data frame - we convert it to a tibble.
    
```{r setup, echo = FALSE, message = FALSE}
library(ISLR)
library(tidyverse)

college <- as_tibble(College)
```

## c) - College Data

### i)

```{r college_summary}
summary(college)
```

### ii)

Create a scatterplot matrix of the first ten variables.	

```{r college_ggpairs, echo = FALSE, message = FALSE}
	library(GGally)
	college %>% ggpairs(columns = 1:10)
```

### iii)

Create a boxplot of *Outstate* versus *Private*

```{r college_boxplot}
opts_knit$set(base.uri='/')
college %>% ggplot + geom_boxplot(aes(Private, Outstate))
```

### iv)
Create a new variable *Elite* by binning the *Top10perc* variable. We divide universities into two groups depending on whether or not the proportion of students coming from the top 10% of their high schools exceeds 10%.

```{r elite_mutation}
college %>% mutate(Elite = factor(ifelse(Top10perc > 50, "Yes", "No")))
```
We now get a summary of the *Elite* 	

```{r elite_summary}
summary(college['Elite'])
```

### v)

Create some histograms with differing bins for the quantitative variables:

* The amount of enrollments in bin widths of 100

```{r enrollment_histogram}
	college %>% ggplot + geom_histogram(aes(Enroll), binwidth = 100)
```

* The number of PhD students in bin widths of 1

```{r phd_histogram}
college %>% ggplot + geom_histogram(aes(PhD), binwidth = 1)
```

# 9 - Auto data

###	a) Which predictors are quantative and which are qualitative?

 **Quantitative** - mpg, displacement, horsepower, weight, acceleration.
 **Qualitative** - cylinders, year, origin, name.

### b) What is the range of each quantitative predictor?

```{r mtcars_quantitative_range}
auto <- as.tibble(mtcars)
auto %>% summarise_at(vars(mpg, disp, hp, wt), funs(min, max))
```

### c) What is the mean and standard deviation for each quantitative predictor?

```{r mtcars_quantitative_mean_sd}
auto %>% summarise_at(vars(mpg, disp, hp, wt), funs(mean, sd))
```

### d) Remove the 10th to 85th observations. What is the range, mean and standard deviation of each predictor?

```{r mtcars_quantitative_removed}
auto[-c(10:85),] %>% summarise_at(vars(mpg, disp, hp, wt), funs(mean, sd))
```

### e) Some graphical representations

Relationship between the miles per gallon and the horsepower, with the colour of the points representing the number of cylinders.

```{r mtcars_hp_vs_mpg}
auto %>% ggplot + geom_point(aes(hp, mpg, colour = as.factor(cyl)))
```

### f) What variables appear to be useful in predicting mpg?

Using `ggpairs` we can see a relationship between *mpg* and *disp*, *hp* and *wt*.
	


