# Chapter 3 - Applied

## 8) Auto data set

```{r applied_setup, echo = F, message = F}
library(tidyverse)
library(broom)
library(ISLR)
```

### a)
*Use the `lm()` function to perform a simple linear regression with `mpg` as the response and `horsepower` as the predictor. Print the results and comment on the output.*

```{r applied_auto_linear}
auto <- as_tibble(Auto)
lm_auto <- auto %>% lm(mpg ~ horsepower, .)
lm_auto %>% tidy()
```

#### i) 
*Is there a relationship between the predictor and the response?*

To determine if there is a relationship, we need to look at two items: the p-values for the coefficients, the F-statistic, and the p-value for the F-statistic:
```{r applied_p_val_and_f_stat}
lm_auto %>% tidy()
lm_auto %>% glance()

```

We in the `tidy()` output we can see the p-value for the intercept and the coefficient is very small, indicating a very low probability for the null hypothesis.A

In the `glance()` output, we see the `statistic` column (the F-statistic) is high at around 600, with a small p-value. This indicates that there is a relationship.

#### ii)
*How strong is the relationship*

To test how string the relationship is, we can look at the residual standard error (RSE) and the R^2 value.

We look at the `glance()` output again and see the R^2 value is .606. Recall that the R^2 is a value between 0 and 1 that is the 'proportion of the variance explained'. We can consider the relationship reasonably strong;

#### iii)
*Is the relationship positive or negative?*
There is a negative relationship between `mpg` and `horsepower`, thus miles per gallon goes down as horsepower goes up. This aligns with our conceptual idea.

#### iv) 
*What is the predicted `mpg` associated with a `horsepower` of 98? What are the associated 95% confidence and prediction intervals?*

We use the `interval` argument to the predict function. Note that the default `level` argument of `predict()` is 0.95 (95%).
```{r applied_linear_prediction}
predict(lm_auto, tibble(horsepower = 98), interval = 'confidence')
predict(lm_auto, tibble(horsepower = 98), interval = 'predict')
```

### b)
*Plot the response and the predictor. Display the least squares regression line.*

```{r applied_auto_mpg_hp}
auto %>% ggplot(aes(mpg, horsepower)) + geom_point() + geom_smooth(method = 'lm', formula = 'mpg ~ horsepower')
```

### c) 
*Produce diagnostic plots (Resid v Leverage, Resid v Fitted, Fitted v Std Resid) and comment on any problems*

First off, lets have a look at the residuals versus the leverage:
```{r applied_auto_resid_v_fitted}
augment(lm_auto) %>% ggplot(aes(.hat, .resid)) + geom_point()
```

Now we look at the fitted versus the residuals:
```{r applied_auto_fitted_v_residuals}
augment(lm_auto) %>% ggplot(aes(.fitted, .resid)) + geom_point()
```





