# Chapter 3 - Applied

## 8) Auto data set

```{r applied_setup, echo = F, message = F}
library(tidyverse)
library(broom)
library(ISLR)
```

### a)
*Use the `lm()` function to perform a simple linear regression with `mpg` as the response and `horsepower` as the predictor. Print the results and comment on the output.*

```{r applied_auto_linear}
auto <- as_tibble(Auto)
lm_auto <- auto %>% lm(mpg ~ horsepower, .)
lm_auto %>% tidy()
```

#### i) 
*Is there a relationship between the predictor and the response?*

To determine if there is a relationship, we need to look at two items: the p-values for the coefficients, the F-statistic, and the p-value for the F-statistic:
```{r applied_p_val_and_f_stat}
lm_auto %>% tidy()
lm_auto %>% glance()

```

We in the `tidy()` output we can see the p-value for the intercept and the coefficient is very small, indicating a very low probability for the null hypothesis.A

In the `glance()` output, we see the `statistic` column (the F-statistic) is high at around 600, with a small p-value. This indicates that there is a relationship.

#### ii)
*How strong is the relationship*

To test how string the relationship is, we can look at the residual standard error (RSE) and the R^2 value.

We look at the `glance()` output again and see the R^2 value is .606. Recall that the R^2 is a value between 0 and 1 that is the 'proportion of the variance explained'. We can consider the relationship reasonably strong;

#### iii)
*Is the relationship positive or negative?*
There is a negative relationship between `mpg` and `horsepower`, thus miles per gallon goes down as horsepower goes up. This aligns with our conceptual idea.

#### iv) 
*What is the predicted `mpg` associated with a `horsepower` of 98? What are the associated 95% confidence and prediction intervals?*

We use the `interval` argument to the predict function. Note that the default `level` argument of `predict()` is 0.95 (95%).
```{r applied_linear_prediction}
predict(lm_auto, tibble(horsepower = 98), interval = 'confidence')
predict(lm_auto, tibble(horsepower = 98), interval = 'predict')
```

### b)
*Plot the response and the predictor. Display the least squares regression line.*

```{r applied_auto_mpg_hp}
auto %>% ggplot(aes(mpg, horsepower)) + geom_point() + geom_smooth(method = 'lm', formula = 'y ~ x')
```

### c) 
*Produce diagnostic plots (Resid v Leverage, Resid v Fitted, Fitted v Std Resid) and comment on any problems*

First off, lets have a look at the residuals versus the leverage:
```{r applied_auto_leverage_v_fitted}
augment(lm_auto) %>% ggplot(aes(.hat, .resid)) + geom_point()
```

There are a few points up in the top right. We take a look at the Cook's distance for the observations.

```{r applied_auto_cooks_distance}
augment(lm_auto) %>% mutate(i = 1:n()) %>% ggplot(aes(i, .cooksd)) + geom_bar(stat = 'identity')
```

A couple of high points but all below 1.

Now we look at the fitted versus the residuals, and also fit a quadratic regression. We see a bit of a U shape, indicating potential non-linearity in the data.
```{r applied_auto_fitted_v_residuals}
augment(lm_auto) %>% ggplot(aes(.fitted, .resid)) + geom_point() + geom_smooth(method = 'lm', formula = 'y ~ poly(x,2)')
```

## 9) Multiple Linear Regression - Auto Data Set

```{r applied_mult_auto_setup, message = F}
library(GGally)
library(corrplot)
```

### a)
*Produce a scatterplot matrix which includes all the data in the data set*
```{r applied_mult_auto_pairs, message = F}
auto %>% select(-'name') %>% ggpairs()
```

### b)
*Compute the matrix of correlations between the variables.*
```{r applied_mult_auto_corr}
auto %>% select(-'name') %>% cor() %>% corrplot(method = 'color')
```

### c)
*Perform a multiple linear regression with `mpg` as the response and all other variables except `name` as the predictors.*

```{r applied_mult_auto_lm}
lin_reg_auto <- lm(mpg ~ . -name, auto)
tidy(lin_reg_auto)
glance(lin_reg_auto)
```
#### i)
*Is there a relationship between the predictors and the response?*
We test the null hypothesis of "are all of the regression coefficients zero?". The F-statistic 252 (far greater than 1) and has a p-value of 2e-139, indicating a low probability that this is just by chance. We can therefore say there is a relationship between the predictors and the response.

#### ii)
We look at the p-values for each of the predictors. The predictors which have a high probability of having an effect on the `mpg`, holding all others constant, appear to be `weight`, `year`, `displacement` and `origin`.

#### iii)
*What does the coefficient for the `year` variable suggest?*
The year coefficient suggests that a cars `mpg` gets larger - and therefore better - the later a car was made.

### d)
*Plot the diagnostic plots of the regression and comment on any problems with the fit.*

We go through our usual plots - first off is looking at the residuals versus the leverage:
```{r applied_multi_auto_leverage}
augment(lin_reg_auto) %>% mutate(i = 1:n()) %>% ggplot(aes(i, .cooksd)) + geom_bar(stat = 'identity')
```

We don't see values with a significant Cook's distance. We move on to the fitted versus the residuals:
```{r applied_auto_fitted_resid}
augment(lin_reg_auto) %>% ggplot(aes(.fitted, .resid)) + geom_point() + geom_smooth(method = 'lm', formula = 'y~poly(x,2)')
```
There is some evidence of the non-linearity of the results.

### e) 
*Use the `*` and `:` symbols to fit linear regressions with interaction effects. Are any interactions statistically significant?*
A `*` adds the predictors and the interaction term, whereas the `:` only adds the interaction term. I.e. `x\*y == x + y + x:y`.

Let's have a think about potential interactions - I think weight and year could interact, given the changes in materials. There could also be an f,Let's have a think about potential interactions - I think weight and year could interact, given the changes in materials. There could also be and interaction between cylinders and displacement:

```{r applied_auto_interaction}
lm(mpg ~ weight*year + cylinders*displacement, auto) %>% tidy()
```

All of the values appear to be reasonably statistically significant. In fact, if we have a look at the fitted vs residuals, it looks much better than before:
```{r applied_auto_interaction_fitted}
lm(mpg ~ weight*year + cylinders*displacement, auto) %>% augment() %>% ggplot(aes(.fitted, .resid)) + geom_point() + geom_smooth()
```

### f)
*Try different transformations of the variables and comment ont the findings.*

We try a few different transformations and pipe them through to the fitted versus residuals graph:

```{r applied_auto_transformations}
lm(mpg ~ sqrt(horsepower), auto) %>% augment() %>% ggplot() + geom_point(aes(.fitted, .resid))
lm(mpg ~ log(horsepower), auto) %>% augment() %>% ggplot() + geom_point(aes(.fitted, .resid))
lm(1/mpg ~ horsepower, auto) %>% augment() %>% ggplot() + geom_point(aes(.fitted, .resid))
lm(1/mpg ~ horsepower + weight*year, auto) %>% augment() %>% ggplot() + geom_point(aes(.fitted, .resid))
lm(1/mpg ~ horsepower + weight*year, auto) %>% glance()
```

The last one looks quite good.


## 10) Carseats Data Set

### a) 
*Fit a multiple regression model to predict `Sales` using `Price`, `Urban`, and `US`.*

```{r applied_carseats_10_a}
carseats <- as_tibble(Carseats)
cs_regress <- lm(Sales ~ Price + Urban + US, carseats)
cs_regress %>% tidy()
```

### b)
*Provide an interpretation of each coefficient in the model.*
* (Intercept) - the average number of sales of carseats, ignoring all other factors. 
* Price - the regression indicates a relationship between price and sales, given the low p-value of the t-statistic. An increase in price of a dollar results in a decrease of 54 carseats solds. 
* UrbanYes - given the high p-value, there doesn't appear to be a relationship between sales and whether a store is urban.
* USYes - given the low p-value, the store bein in the US results in 1200 more carseats being sold.

### c)
*Write out the model in equation form, being careful to handle the qualitative variables properly.*
`Sales = x * Price + y * Urban + z * US, where [Urban = Yes => y = 1|Urban = No =>  y = 0] & [US = Yes => z = 1|US = No => z = 0]`

### d)
*For which of the predictors can you reject the null hypothesis H 0 : βj = 0?*
The null hypothesis can be rejected for `Price` and `US`.

### e)
*On the basis of your response to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the outcome.*
```{r applied_carseats_10_e}
cs_regress_reduced <- lm(Sales ~ Price + US, carseats)
cs_regress_reduced %>% tidy()
```

### f)
*How well do the models in (a) and (e) fit the data?*

We look at the F-statistic and its assoicated p-value to determine how well the models fit:
```{r applied_carseats_10_f}
cs_regress %>% glance()
cs_regress_reduced %>% glance()
```

We see no increase in the R-sqaured value - but we do see an increase in the F-statistic, and a decrease in its p-value.

### g)
*Using the model from (e), obtain 95 % confidence intervals for the coefficient(s).*

```{r applied_carseats_10_g}
cs_regress_reduced %>% confint()
```

### h)
*Is there evidence of outliers or high leverage observations in the model from (e)?*

```{r applied_carseats_10_h}
cs_regress_reduced %>% augment() %>% ggplot(aes(.fitted, .resid)) + geom_point()
cs_regress_reduced %>% augment() %>% ggplot(aes(.hat, .resid)) + geom_point()
```

There doesn't appear to be any non-linearity in the daata, and the high leverage points don't appear to affect the data substantially.

## 11) T-statistic and null-hypothesis
*In this problem we will investigate the t-statistic for the null hypothesis H 0 : β = 0 in simple linear regression without an intercept. To begin, we generate a predictor x and a response y*

```{r applied_11}
set.seed(1)
linear_response <- tibble(x = rnorm(100), y = 2 * x + rnorm(100))
linear_response %>% ggplot(aes(x,y)) + geom_point()
```

### a) 
* Perform a simple linear regression of y onto x , without an intercept. Report the coefficient estimate β̂, the standard error of this coefficient estimate, and the t-statistic and p-value associated with the null hypothesis H 0 : β = 0. Comment on these results.*

```{r applied_11_a}
lr_reg <- lm(y ~ x + 0, linear_response)
lr_reg %>% tidy()
```

The coefficient estimate is 1.99 - very close to the 2 that we used to generate the points. The standard error is .106, the t-statistic is 18.7 and the p-value is 2.64e-34. The t-statistic tells us how many standard deviations the coefficient is away from 0. The p-value gives us the probability that the null hypothesis - that the coefficient is 0 - is true.

### b) 
*Now perform a simple linear regression of x onto y without an intercept, and report on the same metrics as a)*

```{r applied_11_b}
lr_reg_reverse <- lm(x ~ y + 0, linear_response)
lr_reg_reverse %>% tidy()
```

The coefficient estimate and the standard error have changed, which is as expected. The t-statistic and p-value remain the same.

### c)
*What is the relationship between the results obtained in (a) and (b)?*

It's the same line, thus the overall results obtained are the same.

## 12) Simple Linear Regression

### a) 
*Recall that the coefficient estimate β̂ for the linear regression of Y onto X without an intercept is given by (3.38). Under what circumstance is the coefficient estimate for the regression of X onto Y the same as the coefficient estimate for the regression of Y onto X?*

When the sum of the squares of the y observations is the same as the sum or the squares of the x observations.

### b)
Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is different from the coefficient estimate for the regression of Y onto X.

```{r applied_12_b}
set.seed(1)
tibble(x = rnorm(100), y = 4*x) %>% lm(y ~ x, .) %>% tidy()
set.seed(1)
tibble(x = rnorm(100), y = 4*x) %>% mutate(y = sample(
```

### c)
*Generate an example in R with n = 100 observations in which the coefficient estimate for the regression of X onto Y is the Asame as the coefficient estimate for the regression of Y onto X.*

To have the same coefficients, we need the sum of the squares to be the same. To do this, we generate 100 random values for `X`, and we use the same values for `Y`. However we re-order the values so we don't simply get a `y = x` function. We can use the `sample_n()` function for this:

```{r applied_12_c}
set.seed(1)
sample_data <- tibble(x = rnorm(100)) %>% 
    mutate(y = (sample_n(., 100) %>% .[['x']]))
lm(y ~ x, sample_data) %>% tidy()
lm(x ~ y, sample_data) %>% tidy()
```


