---
title: "Chapter 3 - Linear Regression - Lab Exercises"
author: "Greg Foletta"
output:
  html_document:
    theme: cerulean
    highlight: tango
---

## Simple Linear Regression

Load the required libraries, and convert the the Boston data to a tibble:
```{r setup, message = FALSE}
library(tidyverse)
library(broom)
library(MASS)

boston <- as.tibble(Boston)
```

Calculate the linear regression of lstat (lower status of population) on to medv (median value of owner-occupied homes).

```{r}    
lm_boston <- lm(medv ~ lstat, boston)

lm_boston

summary(lm_boston)
confint(lm_boston)
```

We can extract some of the key variables of the regression using the `augment()` function:
```{r}
augment(lm_boston)
```

Using this, we can generate a plot with:

* The fitted line using `geom_smooth()`.
* The (lstat, medv) x/y points.
* The predicted value of medv given the linear regression.
* Segments linking the actual and predicted values of medv.

```{r medv_lstat_plot}
augment(lm_boston) %>% ggplot(aes(lstat, medv)) + 
    geom_smooth(method = 'lm') + 
    geom_point(alpha = .4) + 
    geom_point(aes(lstat, .fitted), shape = 1) + 
    geom_segment(aes(xend = lstat, yend = .fitted), alpha = .5, colour = 'grey')
```

### Predicted versus Residuals

We plot the predited values versus the residuals to get an idea if the

```{r medv_lstat_predict_versus_residuals}
augment(lm_boston) %>% 
    ggplot(aes(.fitted, .resid)) + 
    geom_point() + 
    geom_smooth()
```

We can see a U shape, indicating non-linearity in the data.

### Leverage

```{r medv_lstat_leverage}
augment(lm_boston) %>% 
    mutate(index = c(1:nrow(augment(lm_boston)))) %>% 
    ggplot(aes(index, .hat)) + 
    geom_point()
```

## Multiple Linear Regression

To perform a linear regression with multiple variables, add the predictors together with `+`:

```{r boston_mult_lm}
mult_lm_boston <- lm(medv ~ lstat + age, data = boston)
summary(mult_lm_boston)
```

To perform a regression against all of the predictors, the `.` can be used:

```{r boston_mult_all_lm}
mult_lm_boston <- lm(medv ~ ., boston)
summary(mult_lm_boston)
```

To perform a regression with all predictors except one:
```{r boston_mult_lm_bar_one}
mult_lm_boston <- lm(medv ~ .-age, boston)
```

### Variance Inflation Factors

```{r vif, message = FALSE}
library(car)
vif(mult_lm_boston) %>% 
    as.tibble() %>% 
    mutate(var = names(vif(mult_lm_boston))) %>% 
    ggplot(aes(var, value)) + 
    geom_bar(stat = "identity")
```

## Interaction Terms

To include an interaction term in `lm()`, you can use `pred_1:pred_2`. To include the predictors and the interaction term you can use `pred_1*pred_2`, which is shorthand for `pred_1 + pred_2 + pred_1:pred2`.

```{r boston_mult_interaction}
mult_lm_boston <- lm(medv ~ lstat*age, boston)
summary(mult_lm_boston)
```

## Non-linear Transformations of Predictors

The `lm()` function can accommodate non-linear transformations of the predictors. For example to perform a polynomial regression of `lstat` on to `medv`:

```{r boston_poly}
poly_lm_boston <- lm(medv ~ lstat + poly(lstat,2), boston)
```


## Qualitative Predictors

```{r carseats_loading}
library(ISLR)
carseats <- as.tibble(Carseats)
```

```{r carseats_regression}
lm_carseats <- carseats %>% lm(data = ., Sales ~ . + Income:Advertising + Price:Age)

tidy(lm_carseats)
```







	


